The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-10-10 JStatSoft_forestinventory_master-concordance.tex
update Andreas 10.10.2017(2)

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 107 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 115:%
1 97 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 213:%
123 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 218:%
1 69 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
129 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 803:%
135}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 804:%
1 92 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1211:%
137 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1216:%
1 35 1 1 2 11 0 1 2 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1268:%
143 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1273:%
1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1522:%
149 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1527:%
1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 18 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1724:%
155 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1729:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1741:%
161 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1746:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1751:%
167 11 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 108 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 116:%
1 97 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 214:%
124 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 219:%
1 68 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
130 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 6 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 803:%
136}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 804:%
1 92 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1211:%
138 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1216:%
1 35 1 1 2 11 0 1 2 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1268:%
144 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1273:%
1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1522:%
150 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1527:%
1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 18 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1724:%
156 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1729:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1741:%
162 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1746:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1751:%
168 11 1}

### 2017-10-10 rnw/Main structure and modules of the Package.Rnw
update Andreas 10.10.2017(2)

%\section{Methods and Structure of the Package}
\section{Methods and Structure of the Package}
\label{sec:str_and_mod}

@@

\begin{figure}[htb]
\centering
\resizebox{0.42\hsize}{!}{\includegraphics{fig/localdensity_final-01.eps}}
\caption{Artificial representation of a local density surface. The spatial distribution of a hypothetical density function for every point in a forested area is represented as a wavy piecewise constant green surface.  Sample plots (white dots) inform the inventorist of the value of the density function at that point.}
\label{fig:inf_pop_apr}
\end{figure}

@@

Small area estimation does not necessarily refer to small spatial areas but rather to areas that contain little or no terrestrial sample. To formulate this mathematically, we want to make an estimate for a subregion $G$ of the entire inventory area $F$ (Fig. \ref{fig:concmphase_and_sae_b}). As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multi-phase estimation can be efficient. However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals. The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and to apply this model to the small area. The potential bias of applying that model in $G$ is then corrected for by using the model residuals derived from that small area. If there are no terrestrial plots in $G$ (i.e. $n_{2,G}=0$), one cannot correct for a potential model bias in $G$ and has to accept a potential bias in the estimator. These are called \textit{synthetic} estimates and despite their potential bias, it is usually still possible to calculate their design-based variance.

>>>>

\section{Methods and Structure of the Package}
\label{sec:str_and_mod}

@@

\begin{figure}[htb]
\centering
\resizebox{0.42\hsize}{!}{\includegraphics{fig/localdensity_final-01.eps}}
\caption{Artificial representation of a local density surface. The spatial distribution of a hypothetical density function for every point in a forested area is represented as a wavy piecewise constant green surface.  Sample plots (white dots) inform the inventorist of the value of the density function at that point. Note that the plateaus of constant $Y(x)$ values here have the shape of squares whereas in reality they are likely to be formed by the intersection of circles around trees.}
\label{fig:inf_pop_apr}
\end{figure}

@@

Small area estimation does not necessarily refer to small spatial areas but rather to areas that contain little or no terrestrial sample. To formulate this mathematically, we want to make an estimate for a subregion $G$ of the entire inventory area $F$ (Fig. \ref{fig:concmphase_and_sae_b}). As the sample size in the small area, $n_{2,G}$, is usually too small to provide sufficient estimation precision, multi-phase estimation can be efficient. However, $n_{2,G}$ may also be too small to justify fitting a separate regression model just for that area because the estimates produce undesirably large confidence intervals. The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and to apply this model to the small area. The potential bias of applying that model in $G$ is then corrected for by using the empirical model residuals derived from that small area. If there are no terrestrial plots in $G$ (i.e. $n_{2,G}=0$), one cannot correct for a potential model bias in $G$ and has to accept a potential bias in the estimator. These are called \textit{synthetic} estimates and despite their potential bias, it is usually still possible to calculate their design-based variance.

### 2017-10-10 rnw/Special Cases and Scenarios.Rnw
update Andreas 10.10.2017(2)

For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e. only a subset $M_{G}(x) < M(x)$ is included in the small area of interest. In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer guaranteed when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (see \textit{extsynth} and \textit{extpsynth} in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}). In this case, it is adviseable to use the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still ensured.

>>>>

For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e. only a subset $M_{G}(x) < M(x)$ is included in the small area of interest. In this case, the zero mean residual property within the small area, $\frac{\sum_{x \in s_{2,G}}M(x)\hat{R}_{c}(x)}{\sum_{x \in s_{2,G}}M(x)}=0$, is no longer guaranteed when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (see \textit{extsynth} and \textit{extpsynth} in Sections \ref{sec:twophase_sae} and \ref{sec:threephase_sae}). In this case, it is adviseable to use the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still ensured.

### 2017-10-10 rnw/Threephase_Estimators and their Application.Rnw
update Andreas 10.10.2017(2)

Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ denotes the residuals of the full model, whereas $\hat{R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ denotes the residuals of the reduced model. The \textbf{external variance} form under linear regression models is defined as

@@

The unbiased \textbf{point estimates} of the \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then correcting for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e. ${R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$, to the point estimate. The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a residual term likewise in the global estimation (Eq. \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).

@@

The \textbf{synthetic} (\textit{synth}) and \textbf{pseudo synthetic estimator} (\textit{psynth}) can be applied if no terrestrial samples are available in the small area, i.e. $n_{2,G}=0$. Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point (Eq. \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (Eq. \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas. The point estimates are again potentially biased since the zero mean residual property $\int_G R(x)dx=0$ of the full model can not be insured within the small area $G$. Also the variance will be small but to the cost of ignoring the model uncertainties. Note that there is again no external variance formula for the synthetic and pseudo synthetic estimation.

@@

The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) insures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e. $\int_F R(x)dx=\int_G R(x)dx=0$. This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0. The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(0)t}(x), I_G^t(x))$. In other words, when extended option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for \textit{all} observations in the input data frame (i.e. $s_0$). The regression coefficients, point estimates and variance estimates are calculated by replacing $\pmb{Z}$ with $\pmb{\mathbb{Z}}$ (and likewise $\pmb{Z}^{(0)}$ with $\pmb{\mathbb{Z}}^{(0)}$) into Eq. \ref{eq:normequ_3p}, \ref{eq:covar3p}, \ref{eq:pest_3p_small_psmall} and \ref{eq:3p_synth_psynth}. Just as in the two-phase case, the resulting point estimates are now unbiased and have an associated g-weight variance that accounts for the variability of the regression coefficents resulting from the random sample $s_2$.

>>>>

Note that $\hat{R}(x)=Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ denotes the empirical residuals of the full model, whereas $\hat{R}_{0}(x)=Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$ denotes the empirical residuals of the reduced model. The \textbf{external variance} form under linear regression models is defined as

@@

The unbiased \textbf{point estimates} of the \textbf{small} (\textit{small}) and \textbf{pseudo small area estimator} (\textit{psmall}) are calculated by applying the globally derived reduced and full regression model coefficients to the small area means of the explanatory variables, and then correcting for a potential model bias in $G$ by adding the small area mean of the full model residuals, i.e. $\hat{R}_{G}(x)=Y_G(x)-\pmb{Z}_G^t(x)\hat{\beta}_{s_2}$, to the point estimate. The difference between the mean $\hat{\bar{\pmb{Z}}}^{(0)}_{1,G}$ and the more precise or exact mean $\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}$ and $\bar{\pmb{Z}}^{(0)}_{0,G}$ is again considered as a residual term likewise in the global estimation (Eq. \ref{eq:reg3p_ex} and \ref{eq:reg3p_nex}).

@@

The \textbf{synthetic} (\textit{synth}) and \textbf{pseudo synthetic estimator} (\textit{psynth}) can be applied if no terrestrial samples are available in the small area, i.e. $n_{2,G}=0$. Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point (Eq. \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (Eq. \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas. The point estimates are again potentially biased since $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$ for the full model residuals can not be insured within small area $G$. Also the variance will be small but to the cost of ignoring the model uncertainties. Note that there is again no external variance formula for the synthetic and pseudo synthetic estimation.

@@

The \textbf{extended synthetic} (\textit{extsynth}) and \textbf{extended pseudo synthetic estimator} (\textit{extpsynth}) insures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e.$\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$. This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0. The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^t(x)=(\pmb{\mathbb{Z}}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$, where $\pmb{\mathbb{Z}}^{(0)t}(x)=(\pmb{Z}^{(0)t}(x), I_G^t(x))$. In other words, when the \textit{extended} option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for \textit{all} observations in the input data frame (i.e. $s_0$). The regression coefficients, point estimates and variance estimates are calculated by replacing $\pmb{Z}$ with $\pmb{\mathbb{Z}}$ (and likewise $\pmb{Z}^{(0)}$ with $\pmb{\mathbb{Z}}^{(0)}$) into Eq. \ref{eq:normequ_3p}, \ref{eq:covar3p}, \ref{eq:pest_3p_small_psmall} and \ref{eq:3p_synth_psynth}. Just as in the two-phase case, the resulting point estimates are now unbiased and have an associated g-weight variance that accounts for the variability of the regression coefficents resulting from the random sample $s_2$.

### 2017-10-10 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 10.10.2017(2)

Note that for \textit{internal} linear models the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction which is why it does not appear in the point estimate. More explanation about how to obain the auxiliary means is given in the next subsection.

@@

In the equations for the point estimates (Eq. \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$. A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the model residuals in $G$. This is called the \textit{bias} or \textit{residual correction} term.

@@

The \textbf{synthetic estimator} (\textit{synth}) and \textbf{pseudo synthetic estimator} \textit{psynth} are commonly applied when no terrestrial sample is available within the small area $G$ (i.e. $n_{2,G}=0$). In this case, the point estimates (Eq. \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively. However, the bias correction using the observed residuals is not applied as was the case in the small and pseudo small area estimator (Eq. \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}). Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e. $\int_G R(x)dx=0$. However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$. Also note that the residual variation can no longer be considered in the g-weight variance (Eq. \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}). Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias. Due to the absence of available residuals in $G$, there is also no external variance form for the synthetic and pseudo synthetic estimator.

@@

The synthetic estimators, \textit{synth} and \textit{psynth}, have attractively compact formulas but come with the downside of potential bias in their point estimates which can make the variances seem deceptively. The \textit{small} and \textit{psmall} estimators overcome this issue by using a bias correction term, i.e. $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$. The motivation behind the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the prediction model residuals over the entire area $F$ and the small area $G$ are by definition both zero at the same time, i.e. $\int_F R(x)dx=\int_G R(x)dx=0$. This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0. Recalling that linear models fitted using OLS have zero mean residuals property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model. The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Eq. \ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}. Note that the package functions internally extend the data set by the indicator variable if the \textit{extsynth} or \textit{extpsynth} estimator is called.

>>>>

Note that for \textit{internal} linear models the mean of the empirical residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction (\textit{zero mean residual property}) which is why it does not appear in the point estimate. More explanation about how to obain the auxiliary means is given in the next subsection.

@@

In the equations for the point estimates (Eq. \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}), we see that the globally derived regression coefficients are applied to the exhaustively or non-exhaustively calculated means of the explanatory variables ($\bar{\pmb{Z}}_G$, $\hat{\bar{\pmb{Z}}}_G$) which are now only based on the first-phase sample $s_{1,G}$ located within small area $G$. A potential bias of the regression model predictions in the small area $G$, due to fitting the regression model with data outside of $G$, is then corrected by adding the mean of the empirical model residuals in $G$. This is called the \textit{bias} or \textit{residual correction} term.

@@

The \textbf{synthetic estimator} (\textit{synth}) and \textbf{pseudo synthetic estimator} \textit{psynth} are commonly applied when no terrestrial sample is available within the small area $G$ (i.e. $n_{2,G}=0$). In this case, the point estimates (Eq. \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) are based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vectors $\bar{\pmb{Z}}_G$ and $\hat{\bar{\pmb{Z}}}_G$ respectively. However, the bias correction using the observed residuals $\hat{R}(x)$ is not applied as was the case in the small and pseudo small area estimator (Eq. \ref{eq:pointest_2p_small} and \ref{eq:pepsmall}). Thus, the (pseudo) synthetic estimator is unbiased \textbf{only} if $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$. However, this assumption is unlikely to be fullfilled in practice resulting in a potentially unobservable design-based bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$. Also note that the residual variation can no longer be considered in the g-weight variance (Eq. \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}). Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias. Due to the absence of available residuals in $G$, there is also no external variance form for the synthetic and pseudo synthetic estimator.

@@

The synthetic estimators, \textit{synth} and \textit{psynth}, have attractively compact formulas but come with the downside of potential bias in their point estimates which can make the variances seem deceptively. The \textit{small} and \textit{psmall} estimators overcome this issue by using a bias correction term, i.e. $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)$. The motivation behind the \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) is to use the same mathematically elegant formulas of the (pseudo) synthetic estimators while at the same time ensuring that the empirical residuals of the prediction model in the entire area $F$ and the small area $G$ are by construction both zero at the same time. This is accomplished by extending the vector of auxiliary information $\pmb{Z}(x)$ by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the sample location $x$ lies inside the target small area $G$ and is otherwise set to 0. Recalling that linear models fitted using OLS have zero mean residual property by construction, this leads to unbiased point estimates and is also true if categorical variables are included in the model. The new \textit{extended} auxiliary vector thus becomes $\pmb{\mathbb{Z}}^t(x)=(\pmb{Z}^t(x),I_G(x))$ and can be used to replace its non-extended counterpart $\pmb{Z}^t(x)$ whereever it is used in Eq. \ref{eq:pest_2p_synth_psynth} and \ref{estvarcovaux_G}. Note that the package functions internally extend the data set by the indicator variable if the \textit{extsynth} or \textit{extpsynth} estimator is called.
